---
title: "Health Economics Analysis with AUS-OA"
author: "AUS-OA Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Health Economics Analysis with AUS-OA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE  # Set to TRUE when actually running the examples
)
library(ausoa)
```

# Health Economics Analysis with AUS-OA

This article demonstrates how to conduct health economics analyses using the AUS-OA microsimulation package, focusing on cost-effectiveness analysis for osteoarthritis interventions.

## Introduction

Health economics analysis is a core capability of AUS-OA, enabling evaluation of cost-effectiveness for different OA interventions. This tutorial covers:

1. Cost calculation methodology
2. Quality-Adjusted Life Year (QALY) assessments
3. Cost-effectiveness analysis
4. Budget impact modeling

## Cost Calculation Framework

AUS-OA implements a comprehensive cost calculation framework that captures various cost components:

```{r cost-framework}
# The cost calculation function
?calculate_costs_fcn

# Example cost configuration
cost_config <- list(
  costs = list(
    # TKA (Total Knee Arthroplasty) costs
    tka_primary = list(
      hospital_stay = list(value = 15000, perspective = "healthcare_system"),
      patient_gap = list(value = 2000, perspective = "patient"),
      annual_review = list(value = 200, perspective = "healthcare_system")
    ),
    # Revision surgery costs  
    tka_revision = list(
      hospital_stay = list(value = 20000, perspective = "healthcare_system"),
      patient_gap = list(value = 2500, perspective = "patient")
    ),
    # OA management costs
    oa_management = list(
      consultation = list(value = 80, perspective = "healthcare_system"),
      medication = list(value = 1200, perspective = "patient"),
      physiotherapy = list(value = 1500, perspective = "patient")
    ),
    # Complications costs
    complications = list(
      dvt = list(value = 3000, perspective = "healthcare_system"),
      pji = list(value = 8000, perspective = "healthcare_system")
    )
  )
)
```

## Example: Cost-Effectiveness Analysis

Let's walk through a complete example of performing a cost-effectiveness analysis:

```{r cea-example}
# Create baseline population data
baseline_pop <- data.frame(
  id = 1:5000,
  age = sample(50:80, 5000, replace = TRUE),
  sex = sample(c(0, 1), 5000, replace = TRUE),
  bmi = rnorm(5000, mean = 30, sd = 6),
  kl_score = sample(2:4, 5000, replace = TRUE, prob = c(0.4, 0.4, 0.2)),
  has_tka = sample(c(0, 1), 5000, replace = TRUE, prob = c(0.85, 0.15)),
  tka_year = sample(2015:2025, 5000, replace = TRUE, prob = rep(1/11, 11)),
  revision_risk = runif(5000, 0.01, 0.05),
  qaly = runif(5000, 0.5, 0.8)
)

# Define intervention: Enhanced rehabilitation program
intervention_params <- list(
  enabled = TRUE,
  interventions = list(
    enhanced_rehab = list(
      type = "qaly_and_cost_modification",
      start_year = 2025,
      end_year = 2035,
      parameters = list(
        qaly_improvement = 0.05,  # 5% improvement in QALY
        cost_reduction = 0.10    # 10% reduction in management costs
      )
    )
  )
)

# Apply intervention to a copy of the population
intervened_pop <- apply_interventions(
  baseline_pop,
  intervention_params,
  2025
)

# Calculate costs for both scenarios
baseline_costs <- calculate_costs_fcn(baseline_pop, cost_config)
intervened_costs <- calculate_costs_fcn(intervened_pop, cost_config)

# Compare outcomes
baseline_qaly <- calculate_qaly(baseline_pop)
intervened_qaly <- calculate_qaly(intervened_pop)

# Calculate incremental cost-effectiveness ratio (ICER)
incremental_cost <- sum(intervened_costs$cycle_cost_total) - sum(baseline_costs$cycle_cost_total)
incremental_qaly <- sum(intervened_qaly$qaly_total) - sum(baseline_qaly$qaly_total)

icer <- incremental_cost / incremental_qaly
cat("ICER (cost per QALY gained):", round(icer, 2), "\n")
```

## Budget Impact Analysis

AUS-OA can also model budget impact of interventions at the population level:

```{r budget-impact}
# Define population for budget impact analysis
state_pop <- data.frame(
  id = 1:100000,
  age = sample(45:85, 100000, replace = TRUE),
  sex = sample(c(0, 1), 100000, replace = TRUE),
  bmi = rnorm(100000, mean = 29, sd = 6),
  kl_score = sample(0:4, 100000, replace = TRUE, prob = c(0.25, 0.2, 0.25, 0.2, 0.1)),
  state_resident = rep(TRUE, 100000)  # All residents of the target state
)

# Define intervention: Early intervention program
early_intervention <- list(
  enabled = TRUE,
  interventions = list(
    early_intervention = list(
      type = "bmi_modification",
      start_year = 2025,
      end_year = 2030,
      parameters = list(
        uptake_rate = 0.3,      # 30% uptake among eligible
        bmi_change = -3.0,      # 3 kg/m² reduction
        impact_duration = 5     # Effects last 5 years
      )
    )
  )
)

# Apply intervention with scaled parameters for budget impact
budget_impact_pop <- apply_interventions(
  state_pop[1:10000, ],  # Sample for computational efficiency
  early_intervention,
  2025
)

# Calculate budget impact
pre_intervention_costs <- calculate_costs_fcn(state_pop[1:10000, ], cost_config)
post_intervention_costs <- calculate_costs_fcn(budget_impact_pop, cost_config)

# Project to full population
budget_impact <- (sum(post_intervention_costs$cycle_cost_total) - 
                 sum(pre_intervention_costs$cycle_cost_total)) * 10  # Scale up by 10

cat("Annual budget impact:", round(budget_impact, 2), "\n")
cat("Budget impact per capita:", round(budget_impact / nrow(state_pop), 4), "\n")
```

## Sensitivity Analysis

For robust decision-making, AUS-OA supports sensitivity analysis:

```{r sensitivity}
# Perform one-way sensitivity on key parameter
sensitivity_analysis <- function(param_values, base_config) {
  results <- data.frame(
    param_value = param_values,
    costs = numeric(length(param_values)),
    qaly = numeric(length(param_values))
  )
  
  for (i in seq_along(param_values)) {
    # Modify configuration parameter
    config_mod <- base_config
    config_mod$costs$tka_primary$hospital_stay$value <- param_values[i]
    
    # Calculate metrics
    costs <- calculate_costs_fcn(baseline_pop[1:1000, ], config_mod)
    qaly_vals <- calculate_qaly(baseline_pop[1:1000, ])
    
    results$costs[i] <- mean(costs$cycle_cost_total)
    results$qaly[i] <- mean(qaly_vals$qaly_total)
  }
  
  return(results)
}

# Vary TKA cost from $12,000 to $20,000
param_range <- seq(12000, 20000, by = 1000)
sens_results <- sensitivity_analysis(param_range, cost_config)

# View results
head(sens_results)
```

## Visualization

Visualizing health economics results is important for communication:

```{r visualization}
# Create cost-effectiveness plane (simplified)
library(ggplot2)

# Example cost-effectiveness scatter plot
ce_data <- data.frame(
  delta_qaly = c(0.02, 0.05, 0.08, 0.03),
  delta_cost = c(500, 1200, 1800, 800),
  scenario = c("Conservative", "Moderate", "Aggressive", "Alternative")
)

ggplot(ce_data, aes(x = delta_qaly, y = delta_cost, color = scenario)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  labs(
    title = "Cost-Effectiveness Plane",
    x = "Incremental QALYs",
    y = "Incremental Cost ($)"
  ) +
  theme_minimal()
```

## Conclusion

This tutorial has shown how AUS-OA can be used for comprehensive health economics analysis of osteoarthritis interventions. The package provides:

- Detailed cost calculation frameworks
- QALY assessment capabilities
- Cost-effectiveness analysis tools
- Budget impact modeling
- Sensitivity analysis functionality

These capabilities enable evidence-based decision making for OA policy and resource allocation. For more details on specific functions, see the package documentation.