---
title: "Advanced Features of AUS-OA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Features of AUS-OA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Advanced Features of AUS-OA

This vignette covers the advanced features of the AUS-OA package, including parallel processing, performance monitoring, and advanced modeling capabilities.

## Parallel Processing

AUS-OA includes parallel processing capabilities for large-scale simulations:

```{r parallel-example, eval=FALSE}
library(ausoa)
library(parallel)
library(foreach)
library(doParallel)

# Set up parallel processing
n_cores <- detectCores() - 1  # Leave 1 core free
cl <- makeCluster(n_cores)
registerDoParallel(cl)

# Example of running multiple scenarios in parallel
scenarios <- list(
  scenario_1 = list(intervention_effect = 0.05),
  scenario_2 = list(intervention_effect = 0.10),
  scenario_3 = list(intervention_effect = 0.15)
)

# Run scenarios in parallel (pseudo-code example)
results <- foreach(i = 1:length(scenarios), .packages = 'ausoa') %dopar% {
  scenario_data <- scenarios[[i]]
  # run_simulation_with_params(scenario_data)
  # return(results)
}

stopCluster(cl)
```

## Performance Monitoring

The package includes tools for performance monitoring:

```{r perf-example, eval=FALSE}
library(bench)

# Benchmark function performance
perf_results <- mark(
  calculate_costs_fcn(test_data, config),
  calculate_qaly(test_data),
  iterations = 5
)

print(perf_results)

# Memory profiling
library(profmem)
mem_profile <- profmem({
  calculate_costs_fcn(test_data, config)
})
print(mem_profile)
```

## Configuration Management

AUS-OA provides sophisticated configuration management:

```{r config-advanced, eval=FALSE}
# Create custom configuration
custom_config <- list(
  simulation = list(
    time_horizon = 30,
    population_size = 5000,
    start_year = 2025
  ),
  costs = list(
    tka_primary = list(
      hospital_stay = list(value = 15000, perspective = "healthcare_system"),
      patient_gap = list(value = 2000, perspective = "patient")
    ),
    tka_revision = list(
      hospital_stay = list(value = 20000, perspective = "healthcare_system"),
      patient_gap = list(value = 2500, perspective = "patient")
    )
  ),
  utilities = list(
    kl0 = 0.85,
    kl1 = 0.80,
    kl2 = 0.72,
    kl3 = 0.65,
    kl4 = 0.55,
    post_tka = 0.78
  )
)

# Save and load custom configuration
save_config(custom_config, "custom_config.yaml")
loaded_config <- load_config("custom_config.yaml")
```

## Advanced Intervention Types

The package supports various intervention types:

```{r interventions-advanced, eval=FALSE}
# Complex intervention with multiple parameters
complex_intervention <- list(
  enabled = TRUE,
  interventions = list(
    dual_intervention = list(
      type = "combined_modification",
      start_year = 2025,
      end_year = 2035,
      target_population = list(
        min_age = 65,
        max_age = 80,
        kl_score_min = 2
      ),
      parameters = list(
        uptake_rate = 0.75,
        effects = list(
          bmi_reduction = -2.5,
          qaly_improvement = 0.03,
          cost_reduction = 0.05
        )
      )
    )
  )
)

# Apply complex intervention
# result <- apply_interventions(population, complex_intervention, 2025)
```

## Memory Optimization

For large datasets, memory optimization functions are available:

```{r memory-opt, eval=FALSE}
# Optimize data table memory usage
library(data.table)
dt <- as.data.table(large_population_data)

# Convert to optimized format
opt_dt <- optimize_dt_memory(dt)
```

## Model Validation and Quality Assurance

```{r validation, eval=FALSE}
# Validate input data
validation_errors <- validate_input_data(test_data, 
  required_columns = c("id", "age", "sex", "bmi"))

if (length(validation_errors) == 0) {
  cat("Data validation passed\n")
} else {
  cat("Validation errors found:\n")
  print(validation_errors)
}

# Check data consistency
consistency_issues <- check_data_consistency(test_data)
if (length(consistency_issues) > 0) {
  cat("Consistency issues found:\n")
  print(consistency_issues)
}
```

## Error Handling and Validation

AUS-OA includes robust error handling:

```{r error-handling, eval=FALSE}
# Safe wrappers with validation
safe_result <- tryCatch({
  calculate_costs_fcn(test_data, config)
}, error = function(e) {
  cat("Error in cost calculation:", e$message, "\n")
  return(NULL)
})

# Alternative with validation
validated_result <- safe_calculate_costs_fcn(test_data, config)
```

## Summary

This vignette demonstrated the advanced features of the AUS-OA package:

- Parallel processing for scalable simulations
- Performance monitoring and memory optimization
- Advanced configuration management
- Complex intervention design
- Comprehensive validation and error handling

These features enable comprehensive analysis of osteoarthritis interventions in Australia with high performance and reliability.