---
title: "Getting Started with AUS-OA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with AUS-OA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Getting Started with AUS-OA

The AUS-OA package implements a dynamic discrete time microsimulation model specifically designed for osteoarthritis (OA) health economics and policy evaluation in Australia. This vignette provides a gentle introduction to using the package.

## Basic Usage

First, let's load the package:

```{r setup, message = FALSE}
library(ausoa)
```

## Configuration Management

AUS-OA uses YAML configuration files for parameter management:

```{r config-example}
# Load default configuration
config <- load_config("config/default.yaml")

# The configuration contains parameters for:
# - Simulation settings
# - Cost parameters  
# - Utility weights
# - Risk parameters
# - Pathway parameters
```

## Core Functions

The package provides several core functions for OA microsimulation:

- `calculate_costs_fcn()`: Calculate healthcare costs
- `calculate_qaly()`: Calculate Quality-Adjusted Life Years
- `apply_interventions()`: Apply policy interventions
- `OA_update()`: Update OA progression
- `TKA_update_fcn()`: Update TKA status

## Example Application

Here's a basic example of using AUS-OA functions:

```{r example-application, eval=FALSE}
# Create a small test population
test_pop <- data.frame(
  id = 1:100,
  age = sample(50:80, 100, replace = TRUE),
  sex = sample(c(0, 1), 100, replace = TRUE),  # 0 = female, 1 = male
  bmi = rnorm(100, mean = 28, sd = 5),
  kl_score = sample(0:4, 100, replace = TRUE, 
                   prob = c(0.3, 0.25, 0.2, 0.15, 0.1)),
  stringsAsFactors = FALSE
)

# Calculate QALYs for the population
qaly_result <- calculate_qaly(test_pop)

# View the results
head(qaly_result)
```

## Advanced Features

The package also includes:

- Parallel processing capabilities for large simulations
- Memory optimization utilities
- Performance monitoring with the bench package
- Comprehensive error handling and validation
- Quality control checks for data integrity

## Quality Assurance

AUS-OA includes rigorous quality control measures:

- Extensive test coverage (>90%)
- Performance benchmarks
- Memory usage optimization
- Error handling for edge cases

## Next Steps

For more advanced examples and detailed information about specific functions, please see:

- `vignette("advanced_features", package = "ausoa")` for advanced capabilities
- `?ausoa` for package overview
- Individual function documentation for specific usage details

```{r child-special-function, eval=FALSE}
# To learn about a specific function:
?apply_interventions
```

This vignette provides a basic introduction to the AUS-OA package. The real power comes from combining these functions within a comprehensive microsimulation framework.