---
title: "Simulation results"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---

## Overview

This file presents statistics for key variables in the AUS-OA model. To navigate to a specific variable, please click on the respective section in the table of contents.

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
folder <- paste0(here("output", "figures"), "/", scenario)
if (file.exists(folder)) {
  print("Folder already exists. Please specify another input file name.")
  print("To overwrite, delete existing output files.")
} else {
  dir.create(folder)
}

knitr::opts_chunk$set(
  fig.path = paste0(folder, "/")
)
```

```{r Plotting_functions}
## Function to plot simulated time series with confidence intervals
plot_mean_CI <-
  function(data, y_axis_title, mean, lower_CI, upper_CI) {
    fig <-
      ggplot(data, aes(x = year)) +
      geom_line(
        aes(y = {{ mean }}, color = sex)
      ) +
      facet_grid(sex ~ age_group, scales = "free_y") +
      theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_x_continuous(breaks = scales::pretty_breaks()) +
      scale_color_manual(values = colors) +
      scale_fill_manual(values = colors) +
      labs(
        title = "",
        x = "Year",
        y = y_axis_title,
        color = "",
        fill = ""
      )

    if (probabilistic == TRUE) {
      fig <-
        fig +
        geom_ribbon(
          data = data,
          aes(ymin = {{ lower_CI }}, ymax = {{ upper_CI }}),
          fill = "blue", alpha = 0.25
        )
    }
    fig
  }

plot_yvar <-
  function(data, yvar, y_axis_title) {
    fig <-
      ggplot(data, aes(x = year)) +
      geom_line(
        aes(y = {{ yvar }}, color = sex)
      ) +
      facet_grid(sex ~ age_group, scales = "free_y") +
      theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_x_continuous(breaks = scales::pretty_breaks()) +
      scale_color_manual(values = colors) +
      scale_fill_manual(values = colors) +
      labs(
        title = "",
        x = "Year",
        y = y_axis_title,
        color = "",
        fill = ""
      )


    fig
  }

f_plot_distribution <-
  function(var, yr, mean, lower_CI, upper_CI) {
    X <-
      Z %>%
      filter(variable == var & year == yr & age_group != "All 45 years and above")

    fig <-
      X %>%
      ggplot(aes(x = age_group, y = {{ mean }})) +
      geom_col(aes(fill = sex), position = position_dodge(0.8), width = 0.7) +
      scale_fill_manual(values = colors) +
      labs(x = "Age group", y = X$description[1], fill = "") +
      facet_wrap(~year, ncol = 2) +
      theme(legend.position = "top")

    if (probabilistic == TRUE) {
      fig <-
        fig +
        geom_errorbar(
          aes(ymin = {{ lower_CI }}, ymax = {{ upper_CI }}, group = sex),
          width = 0.2, position = position_dodge(0.8)
        )
    }
    fig
  }
```

## Simulation settings

The following table displays the model settings used in this simulation round.

```{r}
sim_setup %>%
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Trends over time {.tabset}

### BMI (overweight and obese) {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "bmi_overweight_or_obese")
```

#### Percent

```{r BMI_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (%)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Frequency

```{r BMI_time_trend_freq}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (N)"), N_mean, N_lower_CI, N_upper_CI)
```

### OA (osteoarthritis) {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "oa")
```

#### Percent

```{r OA_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (%)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Frequency

```{r OA_time_trend_freq}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (N)"), N_mean, N_lower_CI, N_upper_CI)
```

### Medication for OA {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "drugoa")
```

#### Percent

```{r OA_drug_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (%)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Frequency

```{r OA_drug_trend_freq}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (N)"), N_mean, N_lower_CI, N_upper_CI)
```

### TKA (total knee replacement) {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "tka")
```

#### Percent

```{r TKA_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (%)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Frequency

```{r TKA_time_trend_freq}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (N)"), N_mean, N_lower_CI, N_upper_CI)
```

### TKA complications {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "comp")
```

#### Percent

```{r TKA_comp_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (%)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Frequency

```{r TKA_comp_time_trend_freq}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (N)"), N_mean, N_lower_CI, N_upper_CI)
```

### Revision 1

```{r}
Z <-
  Model_stats %>%
  filter(variable == "revision1")
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (N)"), N_mean, N_lower_CI, N_upper_CI)
```

### Revision 2

```{r}
Z <-
  Model_stats %>%
  filter(variable == "revision2")
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (N)"), N_mean, N_lower_CI, N_upper_CI)
```

### Quality adjusted life years {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "qaly")
```

#### Average

```{r QALY_time_trend_mean}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Mean)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Total

```{r QALY_time_trend_sum}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Total)"), Sum_mean, Sum_lower_CI, Sum_upper_CI)
```

### OA cost {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "oacost")
```

#### Average

```{r OA_cost_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Mean)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Total

```{r OA_cost_time_trend_sum}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Total)"), Sum_mean, Sum_lower_CI, Sum_upper_CI)
```

### TKA cost {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "tkacost")
```

#### Average

```{r TKA_cost_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Total)"), Sum_mean, Sum_lower_CI, Sum_upper_CI)
```

#### Total

```{r TKA_cost_time_trend_sum}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Total)"), Sum_mean, Sum_lower_CI, Sum_upper_CI)
```

### Revision cost {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "rehabcost")
```

#### Average

```{r Rehab_cost_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Mean)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Total

```{r Rehab_cost_time_trend_sum}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Total)"), Sum_mean, Sum_lower_CI, Sum_upper_CI)
```

### Revision cost {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "revcost")
```

#### Average

```{r Revision_cost_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Mean)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Total

```{r Revision_cost_time_trend_sum}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Total)"), Sum_mean, Sum_lower_CI, Sum_upper_CI)
```

### Total cost {.tabset}

```{r}
Z <-
  Model_stats %>%
  filter(variable == "totalcost")
```

#### Average

```{r Total_cost_time_trend_percent}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Mean)"), Mean_mean, Mean_lower_CI, Mean_upper_CI)
```

#### Total

```{r Total_cost_time_trend_sum}
Z %>% plot_mean_CI(., str_glue("{Z$description[1]} (Total)"), Sum_mean, Sum_lower_CI, Sum_upper_CI)
```

## Distribution {.tabset}

## BMI

```{r}
Z <- Model_stats
for (yr in sort(unique(Z$year))) {
  print(f_plot_distribution("bmi_overweight_or_obese", yr, Mean_mean, Mean_lower_CI, Mean_upper_CI))
}
```

## OA {.tabset}

```{r}
for (yr in sort(unique(Z$year))) {
  print(f_plot_distribution("oa", yr, Mean_mean, Mean_lower_CI, Mean_upper_CI))
}
```

## TKA

```{r}
for (yr in sort(unique(Z$year))) {
  print(f_plot_distribution("tka", yr, Mean_mean, Mean_lower_CI, Mean_upper_CI))
}
```
