---
title: "Supplementary materials"
author: "AUS-OA"
date: "2024-03-11"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---

## Overview

This notebook recreates all supplementary materials for the paper and saves each plot and underlying data in a separate folder.

```{r Setup}
# Markdown options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

# Packages
if (!require("pacman")) install.packages("pacman")
p_load(here, tidyverse, readxl, arrow)
options(dplyr.summarise.inform = FALSE)

# Select input folder
input_file <- here("input", "scenarios", "choose file.txt")
print("PLEASE CHOOSE THE INPUT FILE FROM THE MENU...")
input_file <- choose.files(input_file)
scenario <- gsub(".xlsx$", "", basename(input_file))
scenario <- gsub("ausoa_input_", "", scenario)
sim_setup <-
  read_excel(input_file, sheet = "Simulation inputs") %>%
  rename(
    param = `Base population parameters`,
    spec  = `Value`
  ) %>%
  filter(!is.na(spec))

# Create supplementary materials folder if it does not exist
folder <- paste0(here("output", "supplementary_materials"), "/", scenario)
if (file.exists(folder)) {
  print("Folder already exists. Please specify another input file name.")
  print("To overwrite, delete existing output files.")
} else {
  dir.create(folder)
}

knitr::opts_chunk$set(
  fig.path = paste0(folder, "/")
)

# Graphics options
# Graph options
my_theme <-
  theme_bw(base_family = "serif") +
  theme(
    axis.title.x = element_text(margin = margin(t = 10), size = 14),
    axis.title.y = element_text(margin = margin(r = 10), size = 14),
    axis.text = element_text(size = 12),
    legend.background = element_blank(),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16)
  )
theme_set(my_theme)
colors <- c(
  "#EE3377", "#0077BB", "#009988", "firebrick", "#33BBEE", "#CC3311",
  "darkmagenta", "black", "tan4"
)
excel <- c("#4472C4", "#ED7D31", "#A5A5A5", "#FFC000", "#5B9BD5", "#70AD47", "#264478", "#FF0000", "#FF9900")
```

## 1.1 Synthetic base population

```{r}
hilda_file <-
  here(
    "input", "population",
    sim_setup$spec[sim_setup$param == "Input data for base population"]
  )

base_file <- here("input", "population", "am.parquet")

HILDA <-
  read_csv(hilda_file, show_col_types = FALSE)

BASE <-
  read_parquet(base_file)
```

#### Figure S1: Age/sex distribution, HILDA vs synthetic AUS-OA population

```{r}
fig_name <- "S1_Age_sex_distribution_HILDA_vs_syn"

Z <-
  bind_rows(
    HILDA %>%
      select(sex, age) %>%
      mutate(source = "obs"),
    BASE %>%
      select(sex, age) %>%
      mutate(source = "syn")
  )

p <-
  Z %>%
  ggplot(aes(x = age, fill = source)) +
  geom_histogram(position = "dodge", bins = 20) +
  facet_wrap(~sex) +
  labs(x = "Age", y = "Count", fill = "") +
  scale_color_manual(values = excel)

pg <- ggplot_build(p)$data[[1]] %>% as_tibble()


ggsave(
  here(
    folder,
    paste0(fig_name, ".png")
  ),
  width = 10, height = 10, units = "in", dpi = 300
)


write_csv(Z, here(
  folder,
  paste0("Raw_", fig_name, ".csv")
))
write_csv(Z, here(
  folder,
  paste0("ggplot_", fig_name, ".csv")
))


p
```

#### Figure S2: Sex/BMI distribution, HILDA vs synthetic AUS-OA population, 2013

```{r}
fig_name <- "S2_Sex_bmi_distribution_HILDA_vs_syn"
Z <-
  bind_rows(
    HILDA %>%
      select(sex, bmi) %>%
      mutate(source = "obs"),
    BASE %>%
      select(sex, bmi) %>%
      mutate(source = "syn")
  )

p <-
  Z %>%
  ggplot(aes(x = bmi, fill = source)) +
  geom_histogram(position = "dodge", bins = 20) +
  facet_wrap(~sex) +
  labs(x = "BMI", y = "Count", fill = "") +
  scale_color_manual(values = excel)

pg <- ggplot_build(p)$data[[1]] %>% as_tibble()


ggsave(
  here(
    folder,
    paste0(fig_name, ".png")
  ),
  width = 10, height = 10, units = "in", dpi = 300
)


write_csv(Z, here(
  folder,
  paste0("Raw_", fig_name, ".csv")
))
write_csv(Z, here(
  folder,
  paste0("ggplot_", fig_name, ".csv")
))


p
```

#### Figure S3: Osteoarthritis by age,sex, and BMI, HILDA vs synthetic AUSOA- population 2013

```{r}
fig_name <- "S3_OA_by_age_sex_distribution_HILDA_vs_syn"
Z <-
  bind_rows(
    HILDA %>%
      select(sex, age, oa) %>%
      mutate(source = "obs"),
    BASE %>%
      select(sex, age, oa) %>%
      mutate(source = "syn")
  )

p <-
  Z %>%
  ggplot(aes(x = age, fill = source)) +
  geom_histogram(position = "dodge", bins = 20) +
  facet_grid(sex ~ oa) +
  labs(x = "Age", y = "Count", fill = "") +
  scale_color_manual(values = excel)

pg <- ggplot_build(p)$data[[1]] %>% as_tibble()


ggsave(
  here(
    folder,
    paste0(fig_name, ".png")
  ),
  width = 10, height = 10, units = "in", dpi = 300
)


write_csv(Z, here(
  folder,
  paste0("Raw_", fig_name, ".csv")
))
write_csv(Z, here(
  folder,
  paste0("ggplot_", fig_name, ".csv")
))


p
```

#### Figure S4: Quality of life/sex/osteoarthritis, HILDA vs synthetic AUS-OA population 2013

```{r}
fig_name <- "S4_Quality_of_life_by_oa_sex_HILDA_vs_syn"
Z <-
  bind_rows(
    HILDA %>%
      select(sex, age, sf6d, oa) %>%
      mutate(source = "obs"),
    BASE %>%
      select(sex, age, sf6d, oa) %>%
      mutate(source = "syn")
  ) %>%
  mutate(oa = as_factor(oa))


p <-
  Z %>%
  ggplot(aes(x = source, y = sf6d, fill = source)) +
  facet_grid(sex ~ oa) +
  geom_boxplot() +
  geom_point(position = "jitter", alpha = 0.2, size = 0.1) +
  labs(x = "", fill = "")

pg <- ggplot_build(p)$data[[1]] %>% as_tibble()


ggsave(
  here(
    folder,
    paste0(fig_name, ".png")
  ),
  width = 10, height = 10, units = "in", dpi = 300
)


write_csv(Z, here(
  folder,
  paste0("Raw_", fig_name, ".csv")
))
write_csv(Z, here(
  folder,
  paste0("ggplot_", fig_name, ".csv")
))


p
```

#### Figure S5: Two-way ratios of the propensity mean squared error, HILDA vs synthetic AUS-OA population, 2013

Need to pre-process data

```{r}
p_load(synthpop)
pop_weight <-
  sim_setup$spec[sim_setup$param == "Population scale"] %>%
  as.numeric()

scale_HILDA <-
  sim_setup$spec[sim_setup$param == "Population adjuster"] %>%
  as.numeric()

input_data <-
  sim_setup$spec[sim_setup$param == "Input data for base population"] %>%
  as.character()

input_data <-
  here("input", "population", input_data)

# RUN THE SYNTHETIC POPULATION GENERATOR
p_load(simPop, haven, tidyr, dplyr, party, synthpop, here)
## STATA dataset
synpop_wave <- read.csv(input_data)
synpop_wave <- synpop_wave[, c(
  "age", "sex", "bmi", "oa", "sf6d",
  "phi", "year12", "drugoa", "drugmh", "mhc", "state",
  "ccount", "hhrhid", "hhweight"
)]

# adjust weighting (leaving as is will create full population)
synpop_wave$hhweight <- synpop_wave$hhweight / pop_weight
# calibration - HILDA weights seem to have a shortfall. Use for model.
synpop_wave$hhweight <- scale_HILDA * synpop_wave$hhweight
# testing
# synpop_wave$hhweight <- 10

# format data
synpop_wave$state <- as.factor(synpop_wave$state)
synpop_wave$oa <- as.factor(synpop_wave$oa)
synpop_wave$phi <- as.factor(synpop_wave$phi)
synpop_wave$year12 <- as.factor(synpop_wave$year12)
synpop_wave$drugoa <- as.factor(synpop_wave$drugoa)
synpop_wave$drugmh <- as.factor(synpop_wave$drugmh)
synpop_wave$ccount <- as.factor(synpop_wave$ccount)
synpop_wave$mhc <- as.factor(synpop_wave$mhc)

# SIMULATING DATASET ------------------------------------------------------

## Selecting variable to define population
inp <- specifyInput(
  data = synpop_wave,
  hhid = "hhrhid",
  strata = "state",
  weight = "hhweight"
)

## Generating structure of synthetic population
simPop <- simStructure(inp,
  method = "direct",
  basicHHvars = c("age", "sex")
)

simPop <- simCategorical(simPop,
  additional = c("oa", "phi", "year12", "ccount", "mhc", "drugoa", "drugmh"),
  method = "distribution", nr_cpus = 1
)

#
# regModel = ~ age + sex + bmi

simPop <- simContinuous(simPop,
  additional = "bmi",
  # regModel = regModel,
  upper = 200000,
  nr_cpus = 1
)


simPop <- simContinuous(simPop,
  additional = "sf6d",
  # regModel = regModel,
  upper = 200000,
  equidist = FALSE,
  nr_cpus = 1
)


## Call population
mysim <- data.frame(popData(simPop))
mysim <- mysim[, c(
  "state", "age", "sex", "bmi", "oa", "mhc", "ccount", "sf6d",
  "phi", "year12", "drugoa", "drugmh"
)]
# Check number of NAs
na_count <- sum(is.na(mysim))
print(na_count)
# Either re-run or set to zero
mysim[is.na(mysim)] <- 0
```

```{r}
utility.tables(mysim, synpop_wave, tables = "twoway", nworst = 4, max.scale = 20)
```
